---
title: "CommuniMap Glasgow: Comprehensive Analysis Report"
output: html_document
params:
  data: NA
  title: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tidyr)
library(stringr)
library(knitr)
library(readr)
library(readxl)
library(lubridate)
library(sf)
library(leaflet)
library(DT)
library(rmarkdown)
library(htmltools)

source(file.path("..", "R", "app_config.R"), local = TRUE)
source(file.path("..", "R", "sentiment_helpers.R"), local = TRUE)
```

# CoLab Dashboard Analysis: `r params$title`

**Report Date:** `r Sys.Date()`

**Overview:** This report summarises activity within the selected CoLab. It highlights how reports are distributed across Glasgow, how activity changes over time, and what themes appear in the report text.

---

## 1. Understanding the Sentiment Score

The sentiment score gives a simple summary of the tone of the report text. It is based on the **Bing sentiment lexicon**, which classifies sentiment-bearing words as positive or negative.

#### **Formula**

$$\text{Sentiment Score} = \text{mean of sentence polarity scores within each report}$$

Each sentence is treated as one entry, scored from the mix of positive and negative words within that sentence, and then averaged up to the report level. Repeated photo-caption boilerplate is stripped out first.

#### **How to read it**

- **Positive values:** more positive language overall
- **Around zero:** fairly neutral or mixed language
- **Negative values:** more negative language overall

This should be treated as a broad guide rather than a perfect measure of how people feel.

---

## 2. Activity Over Time

The chart below shows how many reports were submitted each day. This can help highlight busy periods, quiet periods, or sudden spikes in activity.

```{r ts_plot, fig.height=4, fig.width=9}
params$data %>%
  count(CREATED_DATE) %>%
  ggplot(aes(CREATED_DATE, n)) +
  geom_col(fill = "#2c3e50") +
  theme_minimal() +
  labs(x = "Submission Date", y = "Number of Reports")
```

---

## 3. Where Reports Are Concentrated

This figure shows the Intermediate Zones with the highest number of reports in the current selection. It gives a quick picture of which areas are most active.

```{r dist_plot, fig.height=5, fig.width=9}
tmp_dist <- params$data %>%
  count(IZ_LABEL, sort = TRUE) %>%
  head(15)

ggplot(tmp_dist, aes(reorder(IZ_LABEL, n), n)) +
  geom_col(fill = "#18bc9c") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Intermediate Zone", y = "Total Reports")
```

---

## 4. Example Sentences

This table highlights example sentences with the strongest positive and negative sentiment scores.

```{r sentiment_examples}
sent_data <- compute_sentiment_examples(params$data) %>%
  transmute(
    Sentiment = str_to_title(sentiment),
    Score = round(sentence_score, 2),
    Sentence = sentence
  )

if (nrow(sent_data) > 0) {
  kable(sent_data)
} else {
  cat("No sentiment-bearing sentences available.")
}
```

---

## 5. Summary Table

The table below lists the top 20 Intermediate Zones in the selected data, along with the number of reports and a sentiment score where text is available.

```{r table_logic}
sentence_scores <- compute_sentence_sentiment(params$data)

if (nrow(sentence_scores) > 0) {
  s_data <- sentence_scores %>%
    group_by(.report_id, IZ_CODE) %>%
    summarise(Score = mean(sentence_score), .groups = "drop") %>%
    group_by(IZ_CODE) %>%
    summarise(Score = round(mean(Score, na.rm = TRUE), 2), .groups = "drop")
} else {
  s_data <- data.frame(IZ_CODE = character(), Score = numeric())
}

params$data %>%
  count(IZ_CODE, IZ_LABEL, name = "Count") %>%
  left_join(s_data, by = "IZ_CODE") %>%
  arrange(desc(Count)) %>%
  head(20) %>%
  kable()
```
